---
title: "Malaria Weekly Situtaion Update"
subtitle: "Week 29, 2025"
author: "cPHEM, EPHI"
format: 
 #html: default
 docx: default
  #reference-doc: custom-reference-doc.docx
execute: 
  echo: false
  warning: false
editor: visual
---

```{r}
#| label: Data importing and processing 
#| include: false
pacman::p_load(tidyverse, readxl, import, here, janitor, rio, magrittr, esquisse,
               stringdist, reshape2, scales, ggsci, RColorBrewer, skimr, patchwork)

pacman::p_load(
  rio,           # to import data
  here,          # to locate files
  tidyverse,     # to clean, handle, and plot the data (includes ggplot2 package)
  sf,            # to manage spatial data using a Simple Feature format
  tmap,          # to produce simple maps, works for both interactive and static maps
  janitor,       # to clean column names
  #OpenStreetMap, # to add OSM basemap in ggplot map
  spdep ,         # spatial statistics,
  here
)

malaria_dat <- import(here("C:/Users/henok/OneDrive - ITG/Desktop/Database/Malaria Cluster Data.xlsx"), guess_max = 400000)
# malaria_dat <- import(here("/Users/henoktadesse/Library/CloudStorage/OneDrive-ITG/Desktop/Database/Malaria Cluster Data.xlsx"), guess_max = 400000)

colnames(malaria_dat)
eth_adm3_raw <- read_sf(here("ETH shape file 2024/eth_admbnda_adm3_csa_bofedb_2021.shp"))
eth_adm3 <- eth_adm3_raw %>%  janitor::clean_names()

eth_adm1 <- read_sf(here("ETH shape file 2024/eth_admbnda_adm1_csa_bofedb_2021.shp")) %>%  
  janitor::clean_names()

#Change those read as numeric to character
columns_to_exclude <- c(1:7)
malaria_dat %<>% mutate(across(.cols = where(is.character) & !all_of(columns_to_exclude), .fns = as.numeric))
#Chnage the name to match wiht previous data set for downstream script 
malaria_dat  %<>%  dplyr::rename('Total Malaria (Confirmed and Clinical)' = 'Total Malaria Case',
                           'PosMalaria_RDT_or_Microscopy_Mixed_OutP_Cases' = 'PosMalaria_RDT_or_Microscopy_mixed_OutP_Cases2')
###2. Basic data cleaning ################

# cleaning column names
mal_dat <-  malaria_dat %>% janitor::clean_names() %>% 
  rename(total_case = total_malaria_confirmed_and_clinical,
         outp_case = t_malaria_out_p_cases,
         inp_case = t_malaria_in_p_cases,
         death = t_malaria_in_p_deaths,
         pf = pos_malaria_rdt_or_microscopy_pf_out_p_cases,
         pv = pos_malaria_rdt_or_microscopy_pv_out_p_cases,
         mixed = pos_malaria_rdt_or_microscopy_mixed_out_p_cases,
         test = tm_suspected_fever_examined,
         total_confirmed_s = total_confirmed,
         confirm_rate_s = percent_of_confirmed,
         pos_rate_s = test_positivity)
# computing row wise values 

mal_dat <- mal_dat %>% rowwise() %>% mutate(total_confirmed_s = sum(pf,pv,mixed,na.rm = T),
                                            confirm_rate_s = round((total_confirmed_s/total_case)*100, 2),
                                            pos_rate_s = round((total_confirmed_s/test)*100, 2),
                                            inp_rate = round((inp_case/total_case) * 100, 1),
                                            cfr = round((death/total_case)*100, 3)) %>% ungroup()

nat_trend <- mal_dat %>% group_by(year,epi_week) %>% 
  summarise(total_case = sum(total_case, na.rm = T),
            inp_rate_s = round((sum(inp_case, na.rm = T)/sum(total_case, na.rm = T))* 100, 2),
            inp_rate = round(mean(inp_rate[is.finite(inp_rate)], na.rm = T),1),
            cfr = round(mean(cfr[is.finite(cfr)], na.rm = T), 1),
            .groups = "drop")

mal_dat <- mal_dat %>% 
  mutate(region = str_squish(str_to_title(region))) #Trim and convert to lowercase

wkly_dat <- mal_dat %>% select(region, zone, g_name, year, epi_week, total_case,
                               death, test,  total_confirmed_s, pos_rate_s) %>% 
  filter(year == 2025 & epi_week == max(epi_week[year == 2025], na.rm = T)) %>% 
  group_by(g_name) %>% 
  summarise( total_case = sum(total_case, na.rm = T),
             total_conf = sum(total_confirmed_s, na.rm = T),
             death = sum(death, na.rm = T),
             pos_rate = round(sum(total_case, na.rm = T)/sum(test, na.rm = T) * 100, 2),
             .groups = "drop")

pop <- import(here("Pop/Population by GIS woreda.xlsx"))
ref_gname <- import(here("pop/Shape file attribute table (2).csv"))

wkly_dat <- wkly_dat %>% left_join(pop, by =c("g_name" = "Gnames")) %>% janitor::clean_names() 

wkly_dat <- wkly_dat %>% rowwise() %>% 
  mutate(mal_ar = round(total_conf/population * 1000, 2)) %>% 
  ungroup()
wkly_dat %<>% rename(woreda = g_name)

```

```{r}
#| label: Inline values
#| include: true

# Latest week
latest_week <- nat_trend %>% filter(year == 2025) %>% summarise(max(epi_week, na.rm = TRUE)) %>% pull()

# Current week stats
this_year <- mal_dat  %>% filter(year == 2025)
this_week <- mal_dat %>% filter(year == 2025, epi_week == latest_week)
prev_week <- mal_dat %>% filter(year == 2025, epi_week == latest_week - 1)
same_week_last_year <- mal_dat %>% filter(year == 2024, epi_week == latest_week)
case_this_year <- sum(this_year$total_case, na.rm = T)
death_this_year <- sum(this_year$death, na.rm = T)

# Aggregated national values
cases_this_week <- sum(this_week$total_case, na.rm = TRUE)
cases_last_week <- sum(prev_week$total_case, na.rm = TRUE)
cases_last_year <- sum(same_week_last_year$total_case, na.rm = TRUE)

# Compute % change
week_over_week_change <- round((cases_this_week - cases_last_week) / cases_last_week * 100, 1)
yoy_change <- round((cases_this_week - cases_last_year) / cases_last_year * 100, 1)

# Deaths
deaths_this_week <- sum(this_week$death, na.rm = TRUE)

# Inpatient rate
inpatient_this_week <- round((sum(this_week$inp_case, na.rm = TRUE) / sum(this_week$total_case, na.rm = TRUE)) * 100, 1)
inpatient_last_year <- round((sum(same_week_last_year$inp_case, na.rm = TRUE) / sum(same_week_last_year$total_case, na.rm = TRUE)) * 100, 1)
inpatient_change <- round(((inpatient_this_week - inpatient_last_year) / inpatient_last_year) * 100, 1)

# Case fatality rate
cfr_this_week <- round((sum(this_week$death, na.rm = TRUE) / sum(this_week$total_case, na.rm = TRUE)) * 100, 3)
cfr_last_year <- round((sum(same_week_last_year$death, na.rm = TRUE) / sum(same_week_last_year$total_case, na.rm = TRUE)) * 100, 3)
cfr_change <- round(((cfr_this_week - cfr_last_year) / cfr_last_year) * 100, 1)

# Top 5 regions this week
top_regions <- this_week %>%
  group_by(region) %>%
  summarise(cases = sum(total_case, na.rm = TRUE)) %>%
  mutate(pct = round(cases / sum(cases) * 100, 1)) %>%
  arrange(desc(cases)) %>%
  slice(1:5)


top_regions_text <- paste0(top_regions$region, " (", top_regions$pct, "%)", collapse = ", ")

# Create a vector of bulleted text lines for top 5 regions
top_regions_bullets <- paste0("- ", top_regions$region, " (", top_regions$pct, "%)")


# Total confirmed percent
lab_confirm_rate <- round(sum(this_week$total_confirmed_s, na.rm = TRUE) / sum(this_week$total_case, na.rm = TRUE) * 100, 1)

# Pf proportion
pf_rate <- round(sum(this_week$pf, na.rm = TRUE) / sum(this_week$total_confirmed_s, na.rm = TRUE) * 100, 1)

# Number of woredas reporting at least 1 case
woreda_count <- this_week %>% filter(total_case > 0) %>% distinct(g_name) %>% nrow()

#Death reporting woredas and number 

dth_report <- this_week %>% filter(death > 0) %>% 
  group_by(woreda) %>% summarise(death = sum(death,na.rm = T),
                                 .groups = "drop") %>% arrange(desc(death))

dth_text <- paste0(dth_report$woreda, " (", dth_report$death, ")", collapse = ", ")

#Top burden woredas 

top_10_woredas <- wkly_dat %>% 
  arrange(desc(mal_ar)) %>% 
  slice_head(n = 10)

top_woredas_txt <- paste0(top_10_woredas$woreda, " (", top_10_woredas$mal_ar, " per 1000)", collapse = ", ")

# extracting the top woredas in region


top_woreda_rgns <- wkly_dat %>% left_join(ref_gname, by = "woreda") %>% arrange(desc(mal_ar)) %>% 
  slice_head(n = 10) %>% group_by(region) %>% 
  summarise(num = n(),
            .groups = "drop") %>% arrange(desc(num))

top_woreda_rgns_txt <- paste0(top_woreda_rgns$region, " (", top_woreda_rgns$num, ")", collapse = ", ")
```

## Epidemiological situation

Since January 1, 2025, as of Epi Week `{r} latest_week`, a total of **`r scales::comma(case_this_year)` malaria cases** and **`r death_this_year` deaths** were reported. In Epi Week `r latest_week`, a total of **`r scales::comma(cases_this_week)` malaria cases** and **`r deaths_this_week` deaths** were reported. This represents a **`r abs(week_over_week_change)`% `r ifelse(week_over_week_change<0, "decrement", "increment")`** from last week and a **`r abs(yoy_change)`% `r ifelse(yoy_change<0, "decrement", "increment")`** from the same week in 2024(@fig-plot-national-trend). Theworeda level national incidence and regional level trend of malaria are depicted in @fig-plot-inc-map and @fig-fac-rgn-trnd, respectively.

According to Malaria surveillance report in Epi week `r latest_week`, 2025;

-   `r sum(top_regions$pct)`% of cases were from the five highest reporting regions (`r paste(top_regions_text, collapse = "")`)

-   A total of **`r woreda_count` woredas** reported at least one malaria case. Lab confirmation rate was **`r lab_confirm_rate`%**, with *Plasmodium falciparum* accounting for **`r pf_rate`%** of confirmed cases.

```{r}
#| warning: false
#| 
#| label: fig-plot-national-trend
#| fig-cap: "National trend of Malaria cases; 2019 to 2025"
#| fig-width: 8
#| fig-height: 5
#| 


latest_week <- nat_trend %>% filter(year == 2025) %>% summarise(max(epi_week)) %>% pull()
label_string <- nat_trend %>%
  filter(epi_week == latest_week, year %in% 2019:2025) %>%
  arrange(desc(year)) %>%
  mutate(line = paste0(year, ": ", scales::comma(total_case))) %>%
  pull(line) %>%
  paste(collapse = "\n")
label_y <- nat_trend %>%
  filter(year == 2025, epi_week == latest_week) %>%
  pull(total_case)

# starting the plot
ggplot(nat_trend, aes(x = epi_week, y = total_case, colour = factor(year), group = factor(year))) +
  geom_line(size = 1) +
  geom_point(size = 1.25) +
  #geom_vline(xintercept = 26, linetype = "dashed", colour = "grey40") +
  scale_color_manual(values = c(
    `2019` = "#1f77b4",  # blue
    `2020` = "#e377c2",  # orange 
    `2021` = "orange",  # green
    `2022` = "purple",  # purple
    `2023` = "lightblue",  # pink
    `2024` = "#8c564b",  # brown
    `2025` = "#252525"   # grey
  )) +
  labs(
    #title = "National Trend of Malaria Cases",
    #subtitle = paste("2019 to 2025(Week ",latest_week,")"),
    x = "Epi_Week",
    y = "Total Cases",
    colour = "Year"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line.y.left = element_line(size = 0.5, colour = "black"),
    axis.line.x.bottom = element_line(size = 0.5, colour = "black"),
    legend.position = "right",
    panel.grid = element_blank()
  ) + 
  scale_x_continuous(expand = c(0,0),breaks = pretty_breaks(n = 5)) +  # better x-axis scaling
  scale_y_continuous(expand = c(0,0),labels = scales::comma) +  # comma formatting for large y-axis values
   annotate("label",
            x = latest_week + 0.5,          # adjust horizontal position
            y = label_y,                  # align with 2025's value
            label = paste0("Week ", latest_week, " Cases\n", label_string),
            hjust = 0,
            vjust = 1,
            size = 3.5,
            label.size = 0.3,
            fill = "white",
            family = "Arial")
```

```{r}
#| warning: false
#| include: true
#| label: fig-plot-inc-map
#| fig-width: 8
#| fig-height: 5.5
#| fig-dpi: 300
#| fig-cap: "Woreda level Incidence of Malaria"

 

wkly_dat_sf <- wkly_dat %>% left_join(eth_adm3, by = c("woreda" = "adm3_en")) %>% 
  select(woreda, total_case, total_conf, death, pos_rate, mal_ar, population, mal_ar,
         adm3_pcode, adm2_en, adm2_pcode, adm1_en, adm1_pcode, geometry) %>% 
  st_as_sf()

# preapring breaks and labels
max_ar <- ceiling(max(wkly_dat_sf$mal_ar, na.rm = TRUE))
ar_breaks <- seq(0, max_ar, length.out = 6)
ar_labels <- paste0(
  round(ar_breaks[-length(ar_breaks)], 1), " – ", 
  round(ar_breaks[-1], 1)
)



tm_shape(wkly_dat_sf)+
  tm_polygons("mal_ar", 
              # breaks = ar_breaks,
              # labels = ar_labels,
              breaks = c(0,0.1, 5,10, 50, Inf),
              labels = c("0", ">0 - <5", "5 - <10", "10 - <50", ">=50"),
              palette = c("lightblue2",
                          "#fee391",  
                          "#fdb863",  
                          "#f03b20",
                          "#bd0026" ),
              title = "Malaria AR per 1000 \nCategory",
              border.col = NA) +
  tm_shape(eth_adm1) +
  tm_borders(col = "black", 
             lwd = 1.5) +
  tm_layout(title = paste(" Malaria Incidence Week ", latest_week," 2025"),
            title.position = c("center","top"),
            inner.margins = c(0.05, 0.1, 0.1, 0.02),
            # inner.margins = c(0.25, 0.25, 0.1, 0.02),# bottom, right, top, left margins for content adjustment
            legend.position = c("left" , "bottom"),
            legend.title.size = 1.0,
            legend.text.size = 0.7, 
            fontfamily = "Arial",
            bg.color = "White",
            frame = TRUE
  ) +
  tm_compass(position = c("right", "top"),
             type = "8star", #compass type
             size = 2 #compass size)
  ) +
  tm_scale_bar(
    position = c("right", "bottom"),
    text.size = 0.7 #Adjust scale bar text size
  )


```

```{r}
#| warning: false
#| include: true
#| label: fig-fac-rgn-trnd
#| fig-width: 10
#| fig-height: 6.5
#| fig-dpi: 600
#| fig-cap: "Regional trend of Malaria"

# generating regional level data

rgn_trnd <- mal_dat %>% group_by(region, year, epi_week) %>% 
  summarise(total_case = sum(total_case, na.rm = T),
            total_confirmed_s = sum(total_confirmed_s, na.rm = T),
            test = sum(test, na.rm = T),
            death = sum(death, na.rm = T),
            pf = sum(pf, na.rm = T),
            pv = sum(pv, na.rm = T),
            mixed = sum(mixed, na.rm = T),
            inp_case = sum(inp_case, na.rm = T),
            pos_rate_s = round(sum(total_confirmed_s, na.rm = T)/sum(test, na.rm = T)* 100,2),
            inp_rate = round(sum(inp_case)/sum(total_case)*100,2),
            .groups = "drop")



ggplot(rgn_trnd,aes(x = epi_week, y = total_case, group = year, colour = as.factor(year))) +
  geom_line(linewidth = 1)+
  geom_path(size = 1.25) +
  #geom_vline(xintercept = 26, linetype = "dashed", colour = "blue")+
  facet_wrap(~region, scales = "free_y", ncol = 4)+
  labs( 
    title = "Malaria Case Trend by Region",
    subtitle = paste("2019 to 2025(Week ",latest_week,")"),
    x = "Epidemiological week",
    y = "Total case") +
    scale_color_manual(values = c(
      `2019` = "#1f77b4",  # blue
      `2020` = "#e377c2",  # orange 
      `2021` = "orange",  # green
      `2022` = "purple",  # purple
      `2023` = "lightblue",  # pink
      `2024` = "#8c564b",  # brown
      `2025` = "#252525"   # grey
    )) +
  theme_minimal() +
  theme(
    text = element_text(family = "Arial"),
    strip.text = element_text(size = 12, face = "bold"),  # facet labels style
    strip.background = element_rect(fill = "white", color = "black"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.line.y.left = element_line(size = 1.0, colour = "black"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_blank(),  # remove grid lines
    panel.grid.minor = element_blank()  # Remove minor grid lines for clarity
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 5)) +  # better x-axis scaling
  scale_y_continuous(labels = scales::comma)  # comma formatting for large y-axis values


```

In Epi week `r latest_week`, 2025;

-   The inpatient rate was **`r inpatient_this_week`%,** which is a **`r abs(inpatient_change)`% `r ifelse(inpatient_change < 0, "decrease", "increase")`** from the same week last yea**r (`r inpatient_last_year`%**)( @fig-plot-Inp-trend).

-   The case fatality rate was `r cfr_this_week` which is a **`r abs(cfr_change)`% `r ifelse(cfr_change < 0, "decrease", "increase")`** from the same week last yea**r (`r cfr_last_year`%)**.

-   Death was reported from; `r paste(dth_text, collapse = "")`.

```{r}
#| warning: false
#| 
#| label: fig-plot-Inp-trend
#| fig-cap: "National trend of Malaria Hospitalization rate; 2019 to 2025" 
#| fig-width: 7
#| fig-height: 4

label_string <- nat_trend %>%
  filter(epi_week == latest_week, year %in% 2023:2025) %>%
  arrange(desc(year)) %>%
  mutate(line = paste0(year, ": ", scales::comma(inp_rate_s))) %>%
  pull(line) %>%
  paste(collapse = "\n")
label_y <- nat_trend %>%
  filter(year == 2025, epi_week == latest_week) %>%
  pull(inp_rate_s)

# Starting the figure
nat_trend %>% filter(year > 2022) %>% 
ggplot( aes(x = epi_week, y = round(inp_rate_s,2), colour = factor(year), group = factor(year))) +
  geom_line(size = 1) +
  geom_point(size = 1.25) +
  #geom_vline(xintercept = 26, linetype = "dashed", colour = "grey40") +
  scale_color_manual(values = c(
    `2023` = "#8c564b",  
    `2024` = "#E66C37", 
    `2025` = "#252525"  
  )) +
  scale_y_continuous(labels = scales::comma) +
  labs(
    #title = "National Trend of Malaria Hospitalizaiton rate",
    #subtitle = paste("2019 to 2025(Week ",latest_week,")"),
    x = "Epi_Week",
    y = "Inpatient rate(%)",
    colour = "Year"
  ) +
  theme_minimal(base_family = "Arial") +
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 11, hjust = 0.5),
    axis.title = element_text(size = 12, face = "bold"),
    axis.text = element_text(size = 10),
    axis.line.y.left = element_line(size = 0.5, colour = "black"),
    axis.line.x.bottom = element_line(size = 0.5, colour = "black"),
    legend.position = "right",
    panel.grid = element_blank()
  ) + 
  scale_x_continuous(expand = c(0,0),breaks = pretty_breaks(n = 5)) +   # better x-axis scaling
  annotate("label",
           x = latest_week + 0.5,          # adjust horizontal position
           y = label_y,                  # align with 2025's value
           label = paste0("Week ", latest_week, " Inpatient rate\n", label_string),
           hjust = 0,
           vjust = 0.7,
           size = 3.5,
           label.size = 0.3,
           fill = "white",
           family = "Arial")

```

In Epi week `r latest_week`, 2025,

-   The 10 woredas with the highest incidence of malaria per 1000 people are; `r paste(top_woredas_txt, collapse = "")`(@fig-top-woreda-trend-plot).

-   These highest burden woredas are from ; `r paste(top_woreda_rgns_txt, collapse = "")`

-   As is expected for the minor transmission seasons, which is a major transmission seasons in the southwestern part of the country, **8 out of the 10 highly affected woredas** are from **SWEPR and South Ethiopia Region.**

```{r}
#| warning: false
#| include: false


top_10_woredas <- wkly_dat %>% 
  arrange(desc(mal_ar)) %>% 
  slice_head(n = 10)


top_10_data_24_25 <- mal_dat %>% filter( year > 2023) %>% 
  filter(g_name %in% top_10_woredas$woreda) %>% 
  select(g_name,year, epi_week, total_case) %>% group_by(g_name, year, epi_week) %>% 
  summarise(epiweek_case = sum(total_case, na.rm = TRUE)) %>% 
  ungroup()

# To order the data based on latest week trend 


order_data <- top_10_data_24_25 %>% filter(year == 2025 & epi_week == latest_week) %>% 
  arrange(desc(epiweek_case)) %>% mutate(order = row_number()) 


Top_10_data_24_25_ord <- top_10_data_24_25 %>% left_join(order_data %>%  select(g_name, order),
                                                         by = c("g_name"))

#The trick is making it factor as below to show 

Top_10_data_24_25_ord <- Top_10_data_24_25_ord %>%
  mutate(g_name = factor(g_name, levels = order_data$g_name[order_data$order]))


plot_top_woreda  <- ggplot(Top_10_data_24_25_ord, aes(x = epi_week, y = epiweek_case, group = year, colour = as.factor(year))) +
  geom_line(size = 1) +
  geom_point(size = 1) +
  facet_wrap(~g_name, scales = "free_y", ncol = 4)+
  theme_minimal()+
  labs(title = "Malaria Trend of 10 woredas with highest incidence" ,
       subtitle = paste("2025 as of Week ",latest_week),
       x = "Epi week", y = "Cases") +
  scale_color_manual(values = c("2025" = "black", "2024" = "#E66C37")) +
  theme(
    text = element_text(family = "Arial"),
    strip.text = element_text(size = 12, face = "bold"),  # facet labels style
    strip.background = element_rect(fill = "white", color = "black"),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.subtitle = element_text(size = 12, hjust = 0.5),
    axis.text = element_text(size = 10),
    axis.line.y.left = element_line(size = 1.0, colour = "black"),
    axis.title = element_text(size = 12, face = "bold"),
    panel.grid.major = element_blank(),  # remove grid lines
    panel.grid.minor = element_blank()  # Remove minor grid lines for clarity
  ) +
  scale_x_continuous(breaks = pretty_breaks(n = 5)) +  # better x-axis scaling
  scale_y_continuous(labels = scales::comma)  # comma formatting for large y-axis values


top_10_map <- eth_adm3 %>% filter(adm3_en %in% top_10_woredas$woreda)

map_plot <- tm_shape(eth_adm1) +
  tm_borders() +
  tm_shape(top_10_map) +
  tm_fill(col = "red", alpha = 0.3, title = "Top 10 Woredas") +  # Fill color for the top 10 woredas
  tm_text("adm3_en", size = 0.7, col = "black", auto.placement = TRUE) +  # Label woredas
  tm_borders(lwd = 0.5) +  # Woreda borders
  tm_layout(main.title = paste("Woredas with Highest Malaria Incidence (Week", latest_week,")"),
            main.title.size = 1, 
            main.title.position = "center", 
            inner.margins = c(-0, 0.1, 0, 0.15) ,
            frame = FALSE)
map_plot_gg <- tmap_grob(map_plot)




```

```{r}
#| warning: false
#| label: fig-top-woreda-trend-plot
#| fig-cap: "Trend of Malaria in 10 woredas reporting higest incidence of malaria "
#| fig-width: 16
#| fig-height: 8
 
plot_top_woreda + map_plot_gg + plot_layout(widths = c(1, 1)) 

```

## Response Measures

-   The malaria epidemic response is coordinated, in collaboration with partners, under the national Malaria Incident Management system (NMIMS).

-   A high-level integarted supportive supervision was conducted to regions and higly affected woredas, jointly with the National Malaria Program.

-   Malaria commodities are being distributed to affected regions based on weekly surveillance malaria burden reports.

-   Daily follow-ups are conducted by regions to track surveillance and response efforts at the woreda level, while weekly feedback is provided based on surveillance reports.
